---
VignetteBuilder: knitr
Suggests: BiocStyle, knitr, rmarkdown
title: "Correlation analysis"
author: "Sebastian Schmidt"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmarkdown::html_document:
    theme: cosmo
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: TRUE
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
knitr::opts_knit$set(root.dir = "C:/Users/SebastianSchmidt/OneDrive - Helmholtz Zentrum München/correlation and clustering")

#knitr::opts_knit$set(root.dir = "C:/Users/sschm/OneDrive - Helmholtz Zentrum München/correlation and clustering")


#load libraries

library(knitr)
library(rmarkdown)
library(lattice)
library(Hmisc)

library(DT)
library(gridExtra)
library(kableExtra)
library(plyr)
library(ggplot2)
library(extrafont)
library(readxl)
library(sm)
library(devEMF)
library(tidyr)
library(data.table)
library(nlme)
library(car)
library(lme4)
library(lmerTest)
library(glmmTMB)
library(dplyr)
library(car)
library(tidyverse)
library(ggcorrplot)
library(fdrtool)
library(corrplot)

library(dichromat)
library(RColorBrewer)
library(ggpubr)



```

# Loading of datads

## data sets

First, we load the data sets.

```{r,  include = FALSE, warning = FALSE}
#Load

data <- read_xlsx("all_data_merge.xlsx")
data <- data %>% remove_rownames %>% column_to_rownames(var="Patient")
data_hNPCs <- read_xlsx("all_data_merge_hNPCs.xlsx") 
data_hNPCs <- data_hNPCs %>% remove_rownames %>% column_to_rownames(var="Patient")
data_DAns <- read_xlsx("all_data_merge_DAns.xlsx") 
data_DAns <- data_DAns %>% remove_rownames %>% column_to_rownames(var="Patient")

#Split Ctrl vs sPD
data_Ctrl <- data[data$disease_state == "Ctrl",]
data_sPD <- data[data$disease_state == "sPD",]
```

How does the data set looks like

```{r, echo = FALSE, include = TRUE, warning = FALSE}
datatable(as.data.frame(data), extensions = c('Buttons', 'FixedColumns'), options = list(dom = 'Bfrtip', buttons = c('csv', 'excel'), autoWidth = TRUE, scrollX = TRUE, fixedColumns = list(rightColumns = 1)))
```


#  Calculate Pearson correlations

data set will be filtered. All non-numeric columns will be removed.

```{r, echo = FALSE, include = FALSE, warning = FALSE}

drop <- c("disease_state", "gender")

data <- data[,!colnames(data) %in% drop]
data_sPD <- data_sPD[,!colnames(data_sPD) %in% drop]

```

## Whole dataset {.tabset} 

```{r, echo = FALSE, include = FALSE, warning = FALSE}
##Calculate pearson correlation
cor <- rcorr(as.matrix(data), type = "pearson")

##Extract coefficient and pvalue
cor_coef <- cor$r
cor_pvalue <- cor$P
cor_pvalue[is.na(cor_pvalue)] <- 0

##Flatten correlation matrix
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}

flat <- flattenCorrMatrix(cor$r, cor$P)


##Save tables
write.table(cor_coef, file = "results/tables/correlation_matrix.csv", sep = ",", quote = FALSE, row.names = TRUE)
write.table(cor_pvalue, file = "results/tables/correlation_matrix_pvalue.csv", sep = ",", quote = FALSE, row.names = TRUE)
write.table(flat, file = "results/tables/correlation_matrix_flat.csv", sep = ",", quote = FALSE, row.names = TRUE)

```

### Correlations as table
```{r, echo = FALSE, include = TRUE, warning = FALSE}
datatable(as.data.frame(flat), extensions = c('Buttons', 'FixedColumns'), options = list(dom = 'Bfrtip', buttons = c('csv', 'excel'), autoWidth = TRUE, scrollX = TRUE, fixedColumns = list(rightColumns = 1)))
```

### Correlation plot - circles

```{r,echo = FALSE, fig.width=40, fig.height=40, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
corrplot(cor_coef, 
         col=colorRampPalette(c("darkblue", "white", "red"))(200),
         order = 'hclust', tl.col = 'black'
        ) 

emf("results/figures/correlation_circles.emf", emfPlus= FALSE, coordDPI = 400, 
     width=40, height=40, family= "Arial", 
    pointsize = 30) 
corrplot(cor_coef, 
         col=colorRampPalette(c("darkblue", "white", "red"))(200),
         order = 'hclust', tl.col = 'black'
        ) 
dev.off()
```

### Correlation plot - coefficients
```{r,echo = FALSE, fig.width=40, fig.height=40, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

corrplot(cor_coef, 
         method = "number",
         #type = 'lower', #tl.srt = 45,
         col=colorRampPalette(c("darkblue", "white", "darkred"))(200),
         order = 'hclust', tl.col = 'black'
         
         ) 

emf("results/figures/correlation_coef.emf", emfPlus= FALSE, coordDPI = 400, 
     width=40, height=40, family= "Arial", 
    pointsize = 30) 
corrplot(cor_coef, 
         method = "number",
         #type = 'lower', #tl.srt = 45,
         col=colorRampPalette(c("darkblue", "white", "darkred"))(200),
         order = 'hclust', tl.col = 'black'
         
         ) 
dev.off()



```


## sPD patients only {.tabset} 

```{r, echo = FALSE, include = FALSE, warning = FALSE}
##Calculate pearson correlation
cor_sPD <- rcorr(as.matrix(data_sPD), type = "pearson")

##Extract coeficient and p-value
cor_sPD_coef <- cor_sPD$r
cor_sPD_pvalue <- cor_sPD$P
cor_sPD_pvalue[is.na(cor_sPD_pvalue)] <- 0

##Flatten correlation matrix
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}

flat_sPD <- flattenCorrMatrix(cor_sPD$r, cor_sPD$P)

##Save tables
write.table(cor_sPD_coef, file = "results/tables/correlation_matrix_sPD.csv", sep = ",", quote = FALSE, row.names = TRUE)
write.table(cor_sPD_pvalue, file = "results/tables/correlation_matrix_sPD_pvalue.csv", sep = ",", quote = FALSE, row.names = TRUE)
write.table(flat_sPD, file = "results/tables/correlation_matrix_sPD_flat.csv", sep = ",", quote = FALSE, row.names = TRUE)

```

### Correlations as table
```{r, echo = FALSE, include = TRUE, warning = FALSE}
datatable(as.data.frame(flat_sPD), extensions = c('Buttons', 'FixedColumns'), options = list(dom = 'Bfrtip', buttons = c('csv', 'excel'), autoWidth = TRUE, scrollX = TRUE, fixedColumns = list(rightColumns = 1)))
```

### Correlation plot - circles

```{r,echo = FALSE, fig.width=40, fig.height=40, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
corrplot(cor_sPD_coef, 
         col=colorRampPalette(c("darkblue", "white", "red"))(200),
         order = 'hclust', tl.col = 'black'
                 ) 

emf("results/figures/correlation_sPD_circles.emf", emfPlus= FALSE, coordDPI = 400, 
     width=40, height=40, family= "Arial", 
    pointsize = 30) 
corrplot(cor_sPD_coef, 
         col=colorRampPalette(c("darkblue", "white", "red"))(200),
         order = 'hclust', tl.col = 'black',
         type = 'lower'
        ) 
dev.off()
```

### Correlation plot - coefficients
```{r,echo = FALSE, fig.width=40, fig.height=40, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

corrplot(cor_sPD_coef, 
         method = "number",
         #type = 'lower', #tl.srt = 45,
         col=colorRampPalette(c("darkblue", "white", "darkred"))(200),
         order = 'hclust', tl.col = 'black'
         ) 

emf("results/figures/correlation_sPD_coef.emf", emfPlus= FALSE, coordDPI = 400, 
     width=40, height=40, family= "Arial", 
    pointsize = 30) 
corrplot(cor_sPD_coef, 
         method = "number",
         #type = 'lower', #tl.srt = 45,
         col=colorRampPalette(c("darkblue", "white", "darkred"))(200),
         order = 'hclust', tl.col = 'black',
         type = 'upper',
         ) 
dev.off()



```


# Visualize subset of interesting paramters for sPD {.tabset} 

```{r, echo = FALSE, include = FALSE, warning = FALSE}

col_keep <- read_xlsx("col_names.xlsx")

##Subset corr matrix
keep <- col_keep$use

cor_intPar_sPD_coef <- cor_sPD_coef[colnames(cor_sPD_coef) %in% keep,colnames(cor_sPD_coef) %in% keep]
cor_intPar_sPD_pvalue <- cor_sPD_pvalue[colnames(cor_sPD_pvalue) %in% keep,colnames(cor_sPD_pvalue) %in% keep]

```

## Correlation plot - circles

```{r,echo = FALSE, fig.width=30, fig.height=30, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
corrplot(cor_intPar_sPD_coef, 
         col=colorRampPalette(c("darkblue", "white", "red"))(200),
         order = 'hclust', tl.col = 'black'
        ) 

emf("results/figures/correlation_intPar_sPD_circles.emf", emfPlus= FALSE, coordDPI = 400, 
     width=30, height=30, family= "Arial", 
    pointsize = 30) 
corrplot(cor_intPar_sPD_coef, 
         col=colorRampPalette(c("darkblue", "white", "red"))(200),
         order = 'hclust', tl.col = 'black'
        ) 
dev.off()
```

## Correlation plot - circles; lower

```{r,echo = FALSE, fig.width=30, fig.height=30, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
corrplot(cor_intPar_sPD_coef, 
         col=colorRampPalette(c("darkblue", "white", "red"))(200),
         order = 'hclust', tl.col = 'black',
         type = 'lower'
        ) 

emf("results/figures/correlation_intPar_sPD_circles_lower.emf", emfPlus= FALSE, coordDPI = 400, 
     width=30, height=30, family= "Arial", 
    pointsize = 30) 
corrplot(cor_intPar_sPD_coef, 
         col=colorRampPalette(c("darkblue", "white", "red"))(200),
         order = 'hclust', tl.col = 'black',
         type = 'lower'
        ) 
dev.off()
```

## Correlation plot - circles; upper

```{r,echo = FALSE, fig.width=30, fig.height=30, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
corrplot(cor_intPar_sPD_coef, 
         col=colorRampPalette(c("darkblue", "white", "red"))(200),
         order = 'hclust', tl.col = 'black',
         type = 'upper'
        ) 

emf("results/figures/correlation_intPar_sPD_circles_upper.emf", emfPlus= FALSE, coordDPI = 400, 
     width=30, height=30, family= "Arial", 
    pointsize = 30) 
corrplot(cor_intPar_sPD_coef, 
         col=colorRampPalette(c("darkblue", "white", "red"))(200),
         order = 'hclust', tl.col = 'black',
         type = 'upper'
        ) 
dev.off()
```

## Correlation plot - coefficients
```{r,echo = FALSE, fig.width=30, fig.height=30, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

corrplot(cor_intPar_sPD_coef, 
         method = "number",
         #type = 'lower', #tl.srt = 45,
         col=colorRampPalette(c("darkblue", "white", "darkred"))(200),
         order = 'hclust', tl.col = 'black'
         ) 

emf("results/figures/correlation_intPar_sPD_coef.emf", emfPlus= FALSE, coordDPI = 400, 
     width=30, height=30, family= "Arial", 
    pointsize = 30) 
corrplot(cor_intPar_sPD_coef, 
         method = "number",
         #type = 'lower', #tl.srt = 45,
         col=colorRampPalette(c("darkblue", "white", "darkred"))(200),
         order = 'hclust', tl.col = 'black',
         type = 'lower'
         ) 
dev.off()



```


# Visualize interesting paramters as xy-plot {.tabset} 

```{r, include = FALSE, echo = FALSE, warnings = FALSE}
data <- read_xlsx("all_data_merge.xlsx")
#data <- data %>% remove_rownames %>% column_to_rownames(var="Patient")
```


```{r,echo = FALSE, fig.width=15, fig.height=15, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
a <- ggscatter(data, x = "delta_HY", y = "hNPC_OGDHC_activity",
                add = "reg.line",               
                conf.int = TRUE,                
                color = "disease_state", 
                palette = c("darkblue", "red"),
                add.params = list(color = c("red"), 
                fill = c("red")),
                label = "Patient",
                repel = TRUE,
                shape = "disease_state",
                size = 5
                )+
  stat_cor(aes(color = disease_state))  

b <- ggscatter(data, x = "delta_HY", y = "DAns_OGDHC_activity",
                add = "reg.line",               
                conf.int = TRUE,                
                color = "disease_state", 
                palette = c("darkblue", "red"),
                add.params = list(color = c("red"), 
                fill = c("red")),
                label = "Patient",
                repel = TRUE,
                shape = "disease_state",
                size = 5
                )+
  stat_cor(aes(color = disease_state)) 

c <- ggscatter(data, x = "delta_UPDRS_III", y = "hNPC_OGDHC_activity",
                add = "reg.line",               
                conf.int = TRUE,                
                color = "disease_state", 
                palette = c("darkblue", "red"),
                add.params = list(color = c("red"), 
                fill = c("red")),
                label = "Patient",
                repel = TRUE,
                shape = "disease_state",
                size = 5
                )+
  stat_cor(aes(color = disease_state)) 

d <- ggscatter(data, x = "hNPC_OGDHC_activity", y = "hNPC_YSI_glucose",
                add = "reg.line",               
                conf.int = TRUE,                
                color = "disease_state", 
                palette = c("darkblue", "red"),
                label = "Patient",
                repel = TRUE,
                shape = "disease_state",
                size = 5
                )+
  stat_cor(aes(color = disease_state)) 

e <-  ggscatter(data, x = "delta_HY", y = "hNPC_YSI_glucose",
                add = "reg.line",               
                conf.int = TRUE,                
                color = "disease_state", 
                palette = c("darkblue", "red"),
                add.params = list(color = c("red"), 
                fill = c("red")),
                label = "Patient",
                repel = TRUE,
                shape = "disease_state",
                size = 5
                )+
  stat_cor(aes(color = disease_state)) 

f <- ggscatter(data, x = "hNPC_OGDHC_activity" , y = "DAns_OGDHC_activity",
                add = "reg.line",               
                conf.int = TRUE,                
                color = "disease_state", 
                palette = c("darkblue", "red"),
                label = "Patient",
                repel = TRUE,
                shape = "disease_state",
                size = 5
                )+
  stat_cor(aes(color = disease_state)) 

g <- ggscatter(data, x = "hNPC_OGDHL_protein", y = "hNPC_nf_GLI3_FL_R_ratio",
                add = "reg.line",               
                conf.int = TRUE,                
                color = "disease_state", 
                palette = c("darkblue", "red"),
                label = "Patient",
                repel = TRUE,
                shape = "disease_state",
                size = 5
                )+
  stat_cor(aes(color = disease_state)) 

h <- ggscatter(data, x = "delta_HY", y = "hNPC_OGDHL_protein",
                add = "reg.line",               
                conf.int = TRUE,                
                color = "disease_state", 
                palette = c("darkblue", "red"),
                add.params = list(color = c("red"), 
                fill = c("red")),
                label = "Patient",
                repel = TRUE,
                shape = "disease_state",
                size = 5
                )+
  stat_cor(aes(color = disease_state)) 

i <- ggscatter(data, x = "hNPC_OGDHC_activity" , y = "hNPC_complexI_activity",
                add = "reg.line",               
                conf.int = TRUE,                
                color = "disease_state", 
                palette = c("darkblue", "red"),
                label = "Patient",
                repel = TRUE,
                shape = "disease_state",
                size = 5
                )+
  stat_cor(aes(color = disease_state)) 








figure <- ggarrange(a, b,c,d,e,f,g,h,i + font("x.text", size = 10),
                    ncol = 3, nrow = 3, align = "hv")

figure

emf("results/figures/correlation_sPD_selected_plots.emf", emfPlus= FALSE, coordDPI = 400, 
    width=15, height=15, family= "Arial") 
figure
dev.off()



```

