---
VignetteBuilder: knitr
Suggests: BiocStyle, knitr, rmarkdown
title: "DAns, Mitochondria quantification using mitochondria staining"
author: "Sebastian Schmidt"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmarkdown::html_document:
    theme: cosmo
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: TRUE
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
knitr::opts_knit$set(root.dir = "")
#load libraries

library(knitr)
library(rmarkdown)
library(DT)
library(gridExtra)
library(kableExtra)
library(plyr)
library(ggplot2)
library(extrafont)
library(readxl)
library(sm)
library(devEMF)
library(tidyr)
library(data.table)
library(nlme)
library(car)
library(lme4)
library(lmerTest)
library(glmmTMB)
library(dplyr)
library(car)
library(lme4)


```

# Loading of data

## Data set

First, we load, filter, and merge the data sets.

```{r, echo = FALSE, include = FALSE, warning = FALSE}
#Load
data_nuc <- read.delim("results_Mito/FilteredNuclei_cellbody_final.csv", sep = ",") 
data_cyt <- read.delim("results_Mito/Cytoplasm.csv", sep = ",") 
data_cyt_filtered <- read.delim("results_Mito/Filterd_Cytoplasm.csv", sep = ",") 

#Filter data sets
##FilteredNuclei_cellbody_final
data_nuc$ID_nuc <- paste(data_nuc$ImageNumber, data_nuc$ObjectNumber, sep = "_")
keep <- c("ID_nuc", "ObjectSkeleton_NumberBranchEnds_mitoTracker_Skeleton",	"ObjectSkeleton_NumberBranchEnds_mito_Skeleton",
          "ObjectSkeleton_NumberNonTrunkBranches_mitoTracker_Skeleton", "ObjectSkeleton_NumberNonTrunkBranches_mito_Skeleton",
          "ObjectSkeleton_NumberTrunks_mitoTracker_Skeleton", "ObjectSkeleton_NumberTrunks_mito_Skeleton",
          "ObjectSkeleton_TotalObjectSkeletonLength_mitoTracker_Skeleton", "ObjectSkeleton_TotalObjectSkeletonLength_mito_Skeleton")
data_nuc <- data_nuc[,colnames(data_nuc) %in% keep]

##Cytoplasm
data_cyt$ID_nuc <- paste(data_cyt$ImageNumber, data_cyt$Parent_FilteredNuclei_cellbody_final, sep = "_")
data_cyt$ID_cyt <- paste(data_cyt$ImageNumber, data_cyt$ObjectNumber, sep = "_")
keep <- c("ID_nuc", "ID_cyt" )
data_cyt <- data_cyt[,colnames(data_cyt) %in% keep]

##Filterd_Cytoplasm
data_cyt_filtered$ID_cyt <- paste(data_cyt_filtered$ImageNumber, data_cyt_filtered$Parent_Cytoplasm, sep = "_")
data_cyt_filtered$ID_cyt_filtered <- paste(data_cyt_filtered$ImageNumber, data_cyt_filtered$ObjectNumber, sep = "_")
keep <- c("ID_cyt_filtered", "ID_cyt", "ImageNumber", "ObjectNumber", "Metadata_SampleID", "Metadata_Replicate_ID", "AreaShape_Area", 
          "Mean_Puncta_Mito_AreaShape_Area", "Mean_Puncta_Mito_Intensity_MeanIntensity_Corr_mito", 
          "Mean_Puncta_Mito_Number_Object_Number",
          "Mean_Puncta_mitoTracker_AreaShape_Area", "Mean_Puncta_mitoTracker_Intensity_MeanIntensity_Corr_mitoTracker", 
          "Mean_Puncta_mitoTracker_Number_Object_Number"
          )

data_cyt_filtered <- data_cyt_filtered[,colnames(data_cyt_filtered) %in% keep]


#Merge
data <- merge(data_nuc, data_cyt, by = "ID_nuc")
data <- merge(data, data_cyt_filtered, by = "ID_cyt")


#Sort
sortList <- c("i1JF-R1-018", "iG3G-R1-039", "i1E4-R1-003", "iO3H-R1-005", "i82A-R1-002", "iJ2C-R1-015", "iM89-R1-005", "iC99-R1-007", "iR66-R1-007", "iAY6-R1-003", "iPX7-R1-001", "i88H-R1-002")

data$Metadata_SampleID <- factor(data$Metadata_SampleID, levels = sortList)


#Add disease state information
Ctrl <- c("i1JF-R1-018", "iG3G-R1-039", "i1E4-R1-003", "iO3H-R1-005", "i82A-R1-002")
data$Disease_state <- ifelse(data$Metadata_SampleID %in% Ctrl, "Ctrl", "sPD")


#Add values normalized by cytoplasm area
data$"mitoTracker_MeanArea" <- data$Mean_Puncta_mitoTracker_AreaShape_Area/data$AreaShape_Area
data$"mitoTracker_MeanCount" <- data$Mean_Puncta_mitoTracker_Number_Object_Number/data$AreaShape_Area
data$"mitoTracker_MeanLength" <- data$ObjectSkeleton_TotalObjectSkeletonLength_mitoTracker_Skeleton/data$AreaShape_Area

#Sum Branchpoints
data$Branchpoints <- data$ObjectSkeleton_NumberNonTrunkBranches_mitoTracker_Skeleton + data$ObjectSkeleton_NumberTrunks_mitoTracker_Skeleton

```

How does the data set looks like

```{r, echo = FALSE, include = TRUE, warning = FALSE}
datatable(as.data.frame(data), extensions = c('Buttons', 'FixedColumns'), options = list(dom = 'Bfrtip', buttons = c('csv', 'excel'), autoWidth = TRUE, scrollX = TRUE, fixedColumns = list(rightColumns = 1)))
```

##  Set tresholds {.tabset} 

Applied thresholds are indicated by grey horizontal line.

### Mean_Puncta_mitoTracker_AreaShape_Area

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
ggplot(data,aes(x= Metadata_SampleID, y= Mean_Puncta_mitoTracker_AreaShape_Area, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = 200)
```

### Mean_Puncta_mitoTracker_Number_Object_Number

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
ggplot(data,aes(x= Metadata_SampleID, y= Mean_Puncta_mitoTracker_Number_Object_Number, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic()  + geom_hline(yintercept = 1400)
```

### mitoTracker_MeanArea

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
ggplot(data,aes(x= Metadata_SampleID, y= mitoTracker_MeanArea, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic()  + geom_hline(yintercept = 0.04)
```

### mitoTracker_MeanCount

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
ggplot(data,aes(x= Metadata_SampleID, y= mitoTracker_MeanCount, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic()  + geom_hline(yintercept = 0.45)
```

### mitoTracker_MeanLength

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
ggplot(data,aes(x= Metadata_SampleID, y= mitoTracker_MeanLength, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic()  + geom_hline(yintercept = 0.03)
```

### Branchpoints

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
ggplot(data,aes(x= Metadata_SampleID, y= Branchpoints, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = 40)
```


## Counts per sample

```{r}
#Apply tresholds
data <- subset(data, Mean_Puncta_mitoTracker_AreaShape_Area < 200)
data <- subset(data, Mean_Puncta_mitoTracker_Number_Object_Number < 1400)
data <- subset(data, mitoTracker_MeanArea < 0.04)
data <- subset(data, mitoTracker_MeanCount < 0.45)
data <- subset(data, mitoTracker_MeanLength < 0.03)
data <- subset(data, Branchpoints < 40)

#Save data set
write.csv(data, file = "results_Mito/tables/data_Mito.csv")

```

Cell counts per cell line:
```{r}
#data <- read.csv("results_Mito/tables/data_Mito.csv")
table(data$Metadata_SampleID)
```
Mean cell count:

```{r}
mean(table(data$Metadata_SampleID))
```

# Visualize mitochondrial parameters

Various mitochondrial parameters are visualized for each patient-derived cell line as well as for the disease state
Mean Ctrl levels are indicated by grey horizontal line.

## MitoTracker Area {.tabset} 

### each sample
```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$Mean_Puncta_mitoTracker_AreaShape_Area))

ggplot(data,aes(x= Metadata_SampleID, y= Mean_Puncta_mitoTracker_AreaShape_Area, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-area_violin-plot_patients.svg", width = 20, height = 6, units = "cm")
```

### disease-state
```{r,echo = FALSE, fig.width=8, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$Mean_Puncta_mitoTracker_AreaShape_Area))

ggplot(data,aes(x= Disease_state, y= Mean_Puncta_mitoTracker_AreaShape_Area, group = Disease_state, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-area_violin-plot_Disease-state.svg", width = 8, height = 6, units = "cm")
```

## MitoTracker Count {.tabset} 
### each sample
```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$Mean_Puncta_mitoTracker_Number_Object_Number))

ggplot(data,aes(x= Metadata_SampleID, y= Mean_Puncta_mitoTracker_Number_Object_Number, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-count_violin-plot_patients.svg", width = 20, height = 6, units = "cm")
```

### disease-state
```{r,echo = FALSE, fig.width=8, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$Mean_Puncta_mitoTracker_Number_Object_Number))

ggplot(data,aes(x= Disease_state, y= Mean_Puncta_mitoTracker_Number_Object_Number, group = Disease_state, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-count_violin-plot_Disease-state.svg", width = 8, height = 6, units = "cm")
```

## MitoTracker Intensity {.tabset} 
### each sample
```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$Mean_Puncta_mitoTracker_Intensity_MeanIntensity_Corr_mitoTracker))

ggplot(data,aes(x= Metadata_SampleID, y= Mean_Puncta_mitoTracker_Intensity_MeanIntensity_Corr_mitoTracker, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-intensity_violin-plot_patients.svg", width = 20, height = 6, units = "cm")
```

### disease-state
```{r,echo = FALSE, fig.width=8, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$Mean_Puncta_mitoTracker_Intensity_MeanIntensity_Corr_mitoTracker))

ggplot(data,aes(x= Disease_state, y= Mean_Puncta_mitoTracker_Intensity_MeanIntensity_Corr_mitoTracker, group = Disease_state, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-intensity_violin-plot_Disease-state.svg", width = 8, height = 6, units = "cm")
```

## MitoTracker Mean Area {.tabset} 
### each sample
```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$mitoTracker_MeanArea))

ggplot(data,aes(x= Metadata_SampleID, y= mitoTracker_MeanArea, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-MeanArea_violin-plot_patients.svg", width = 20, height = 6, units = "cm")
```

### disease-state
```{r,echo = FALSE, fig.width=8, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$mitoTracker_MeanArea))

ggplot(data,aes(x= Disease_state, y= mitoTracker_MeanArea, group = Disease_state, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-MeanArea_violin-plot_Disease-state.svg", width = 8, height = 6, units = "cm")
```

## MitoTracker Mean Count {.tabset} 
### each sample
```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$mitoTracker_MeanCount))

ggplot(data,aes(x= Metadata_SampleID, y= mitoTracker_MeanCount, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-MeanCount_violin-plot_patients.svg", width = 20, height = 6, units = "cm")

```

### disease-state
```{r,echo = FALSE, fig.width=8, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$mitoTracker_MeanCount))

ggplot(data,aes(x= Disease_state, y= mitoTracker_MeanCount, group = Disease_state, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-MeanCount_violin-plot_Disease-state.svg", width = 8, height = 6, units = "cm")

```

## MitoTracker Number Branch Ends {.tabset} 
### each sample
```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$ObjectSkeleton_NumberBranchEnds_mitoTracker_Skeleton))

ggplot(data,aes(x= Metadata_SampleID, y= ObjectSkeleton_NumberBranchEnds_mitoTracker_Skeleton, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-BranchEnds_violin-plot_patients.svg", width = 20, height = 6, units = "cm")

```

### disease-state
```{r,echo = FALSE, fig.width=8, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$ObjectSkeleton_NumberBranchEnds_mitoTracker_Skeleton))

ggplot(data,aes(x= Disease_state, y= ObjectSkeleton_NumberBranchEnds_mitoTracker_Skeleton, group = Disease_state, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-BranchEnds_violin-plot_Disease-state.svg", width = 8, height = 6, units = "cm")

```

## MitoTracker Number Branchpoints {.tabset} 
### each sample
```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$Branchpoints))

ggplot(data,aes(x= Metadata_SampleID, y= Branchpoints, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-Branchpoints_violin-plot_patients.svg", width = 20, height = 6, units = "cm")

```

### disease-state
```{r,echo = FALSE, fig.width=8, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$Branchpoints))

ggplot(data,aes(x= Disease_state, y= Branchpoints, group = Disease_state, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-Branchpoints_violin-plot_Disease-state.svg", width = 8, height = 6, units = "cm")

```

## MitoTracker Skeleton Length {.tabset} 
### each sample
```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$ObjectSkeleton_TotalObjectSkeletonLength_mitoTracker_Skeleton))

ggplot(data,aes(x= Metadata_SampleID, y= ObjectSkeleton_TotalObjectSkeletonLength_mitoTracker_Skeleton, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-SkeletonLength_violin-plot_patients.svg", width = 20, height = 6, units = "cm")

```

### disease-state
```{r,echo = FALSE, fig.width=8, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$ObjectSkeleton_TotalObjectSkeletonLength_mitoTracker_Skeleton))

ggplot(data,aes(x= Disease_state, y= ObjectSkeleton_TotalObjectSkeletonLength_mitoTracker_Skeleton, group = Disease_state, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-SkeletonLength_violin-plot_Disease-state.svg", width = 8, height = 6, units = "cm")

```

## MitoTracker Mean Skeleton Length {.tabset} 
### each sample
```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$mitoTracker_MeanLength))

ggplot(data,aes(x= Metadata_SampleID, y= mitoTracker_MeanLength, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-Mean_SkeletonLength_violin-plot_patients.svg", width = 20, height = 6, units = "cm")

```

### disease-state
```{r,echo = FALSE, fig.width=8, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$mitoTracker_MeanLength))

ggplot(data,aes(x= Disease_state, y= mitoTracker_MeanLength, group = Disease_state, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results_Mito/figures/data_Mito-Mean_SkeletonLength_violin-plot_Disease-state.svg", width = 8, height = 6, units = "cm")

```


# Statistical testing using linear mixed effects models {.tabset} 

Nested approach ("Mitochondrial Parameter" ~  Disease_state + (1 | Disease_state:Metadata_SampleID)) to compensate for dependencies within the groups.

## MitoTracker Area

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

mod0 <- lme4::lmer(Mean_Puncta_mitoTracker_AreaShape_Area ~  Disease_state + (1 | Disease_state:Metadata_SampleID), data = data, REML = FALSE); summary(mod0)
Anova(mod0)
```

## MitoTracker Count 

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

mod0 <- lme4::lmer(Mean_Puncta_mitoTracker_Number_Object_Number ~  Disease_state + (1 | Disease_state:Metadata_SampleID), data = data, REML = FALSE); summary(mod0)
Anova(mod0)
```

## MitoTracker Intensity

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

mod0 <- lme4::lmer(Mean_Puncta_mitoTracker_Intensity_MeanIntensity_Corr_mitoTracker ~  Disease_state + (1 | Disease_state:Metadata_SampleID), data = data, REML = FALSE); summary(mod0)
Anova(mod0)
```

## MitoTracker Mean Area

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

mod0 <- lme4::lmer(mitoTracker_MeanArea ~  Disease_state + (1 | Disease_state:Metadata_SampleID), data = data, REML = FALSE); summary(mod0)
Anova(mod0)
```

## MitoTracker Mean Count

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

mod0 <- lme4::lmer(mitoTracker_MeanCount ~  Disease_state + (1 | Disease_state:Metadata_SampleID), data = data, REML = FALSE); summary(mod0)
Anova(mod0)

```

## MitoTracker Number Branch Ends

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

mod0 <- lme4::lmer(ObjectSkeleton_NumberBranchEnds_mitoTracker_Skeleton ~  Disease_state + (1 | Disease_state:Metadata_SampleID), data = data, REML = FALSE); summary(mod0)
Anova(mod0)

```

## MitoTracker Number Branchpoints

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

mod0 <- lme4::lmer(Branchpoints ~  Disease_state + (1 | Disease_state:Metadata_SampleID), data = data, REML = FALSE); summary(mod0)
Anova(mod0)

```

## MitoTracker Skeleton Length

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

mod0 <- lme4::lmer(ObjectSkeleton_TotalObjectSkeletonLength_mitoTracker_Skeleton ~  Disease_state + (1 | Disease_state:Metadata_SampleID), data = data, REML = FALSE); summary(mod0)
Anova(mod0)

```

## MitoTracker Mean Skeleton Length

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

mod0 <- lme4::lmer(mitoTracker_MeanLength ~  Disease_state + (1 | Disease_state:Metadata_SampleID), data = data, REML = FALSE); summary(mod0)
Anova(mod0)

```

