---
VignetteBuilder: knitr
Suggests: BiocStyle, knitr, rmarkdown
title: "NPCs, GLUT3 quantification using antibody staining"
author: "Sebastian Schmidt"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmarkdown::html_document:
    theme: cosmo
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: TRUE
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
knitr::opts_knit$set(root.dir = "")
#load libraries

library(knitr)
library(rmarkdown)
library(DT)
library(gridExtra)
library(kableExtra)
library(plyr)
library(ggplot2)
library(extrafont)
library(readxl)
library(sm)
library(devEMF)
library(tidyr)
library(data.table)
library(nlme)
library(car)
library(lme4)
library(lmerTest)
library(glmmTMB)
library(dplyr)
library(car)
library(lme4)


```

# Loading of data

## Data set

First, we load, filter, and merge the data sets.

```{r, echo = FALSE, include = FALSE, warning = FALSE}
#Load

data <- read.delim("Filterd_Cytoplasm.csv", sep = ",") 

#Filter data sets
keep <- c("FileName_DNA", "ImageNumber", "ObjectNumber", "Metadata_SampleID", "Metadata_Replicate_ID", "AreaShape_Area", "Intensity_MeanIntensity_Corr_GLUT", "RadialDistribution_FracAtD_Corr_GLUT_1of10","RadialDistribution_FracAtD_Corr_GLUT_2of10","RadialDistribution_FracAtD_Corr_GLUT_3of10","RadialDistribution_FracAtD_Corr_GLUT_4of10","RadialDistribution_FracAtD_Corr_GLUT_5of10","RadialDistribution_FracAtD_Corr_GLUT_6of10","RadialDistribution_FracAtD_Corr_GLUT_7of10","RadialDistribution_FracAtD_Corr_GLUT_8of10","RadialDistribution_FracAtD_Corr_GLUT_9of10","RadialDistribution_FracAtD_Corr_GLUT_10of10",
          "RadialDistribution_MeanFrac_Corr_GLUT_1of10","RadialDistribution_MeanFrac_Corr_GLUT_2of10","RadialDistribution_MeanFrac_Corr_GLUT_3of10","RadialDistribution_MeanFrac_Corr_GLUT_4of10","RadialDistribution_MeanFrac_Corr_GLUT_5of10","RadialDistribution_MeanFrac_Corr_GLUT_6of10","RadialDistribution_MeanFrac_Corr_GLUT_7of10","RadialDistribution_MeanFrac_Corr_GLUT_8of10","RadialDistribution_MeanFrac_Corr_GLUT_9of10","RadialDistribution_MeanFrac_Corr_GLUT_10of10"
          )

data <- data[,colnames(data) %in% keep]


#Sort
sortList <- c("i1JF-R1-018", "iG3G-R1-039", "i1E4-R1-003", "iO3H-R1-005", "i82A-R1-002", "iJ2C-R1-015", "iM89-R1-005", "iC99-R1-007", "iR66-R1-007", "iAY6-R1-003", "iPX7-R1-001", "i88H-R1-002")

data$Metadata_SampleID <- factor(data$Metadata_SampleID, levels = sortList)


#Add disease state information
Ctrl <- c("i1JF-R1-018", "iG3G-R1-039", "i1E4-R1-003", "iO3H-R1-005", "i82A-R1-002")
data$Disease_state <- ifelse(data$Metadata_SampleID %in% Ctrl, "Ctrl", "sPD")


```

How does the data set looks like

```{r, echo = FALSE, include = TRUE, warning = FALSE}
datatable(as.data.frame(data), extensions = c('Buttons', 'FixedColumns'), options = list(dom = 'Bfrtip', buttons = c('csv', 'excel'), autoWidth = TRUE, scrollX = TRUE, fixedColumns = list(rightColumns = 1)))
```

##  Set tresholds {.tabset} 

Applied thresholds are indicated by grey horizontal line.


### Area

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
ggplot(data,aes(x= Metadata_SampleID, y= AreaShape_Area, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1) + theme_classic() 
```

### Mean Intensity

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
ggplot(data,aes(x= Metadata_SampleID, y= Intensity_MeanIntensity_Corr_GLUT, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = 0.6)
```



### Intensity cytoplasm ring 10

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
ggplot(data,aes(x= Metadata_SampleID, y= RadialDistribution_FracAtD_Corr_GLUT_10of10, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic()  + geom_hline(yintercept = 0.72)
```

### Intensity cytoplasm ring 9

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
ggplot(data,aes(x= Metadata_SampleID, y= RadialDistribution_FracAtD_Corr_GLUT_9of10, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic()  + geom_hline(yintercept = 0.55) + geom_hline(yintercept = 0.15)
```

### Mean intensity cytoplasm ring 10

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
ggplot(data,aes(x= Metadata_SampleID, y= RadialDistribution_MeanFrac_Corr_GLUT_10of10, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic()  + geom_hline(yintercept = 0.7) + geom_hline(yintercept = 1.2)
```

### Mean intensity cytoplasm ring 9

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
ggplot(data,aes(x= Metadata_SampleID, y= RadialDistribution_MeanFrac_Corr_GLUT_9of10, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic()  + geom_hline(yintercept = 0.85) + geom_hline(yintercept = 1.2)
```




## Counts per sample


```{r, echo = FALSE, include = FALSE, warning = FALSE}
#Apply tresholds
data <- subset(data, Intensity_MeanIntensity_Corr_GLUT < 0.6)
data <- subset(data, RadialDistribution_FracAtD_Corr_GLUT_10of10 < 0.72)
data <- subset(data, RadialDistribution_FracAtD_Corr_GLUT_9of10 < 0.55 & RadialDistribution_FracAtD_Corr_GLUT_9of10 > 0.15)
data <- subset(data, RadialDistribution_MeanFrac_Corr_GLUT_10of10 < 1.2 & RadialDistribution_MeanFrac_Corr_GLUT_10of10 > 0.7)
data <- subset(data, RadialDistribution_MeanFrac_Corr_GLUT_9of10 < 1.2 & RadialDistribution_MeanFrac_Corr_GLUT_9of10 > 0.85)


#Save data set
write.csv(data, file = "results/tables/data_GLUT3.csv")

```


Cell counts per cell line:
```{r}
#data <- read.csv("results/tables/data_GLUT3.csv")
table(data$Metadata_SampleID)
```
Mean cell count:

```{r}
mean(table(data$Metadata_SampleID))
```

# Visualize mitochondrial parameters

Various GLUT parameters are visualized for each patient-derived cell line as well as for the disease state
Mean Ctrl levels are indicated by grey horizontal line.


## Cytoplasmic area {.tabset} 
### each sample
```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$AreaShape_Area))

ggplot(data,aes(x= Metadata_SampleID, y= AreaShape_Area, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results/figures/data_area_violin-plot_patients.svg", width = 20, height = 6, units = "cm")
```

### disease-state
```{r,echo = FALSE, fig.width=8, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$AreaShape_Area))

ggplot(data,aes(x= Disease_state, y= AreaShape_Area, group = Disease_state, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results/figures/data_area_violin-plot_Disease-state.svg", width = 8, height = 6, units = "cm")
```

## Mean intensity {.tabset} 

### each sample
```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$Intensity_MeanIntensity_Corr_GLUT))

ggplot(data,aes(x= Metadata_SampleID, y= Intensity_MeanIntensity_Corr_GLUT, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results/figures/data_mean-intensity_violin-plot_patients.svg", width = 20, height = 6, units = "cm")
```

### disease-state
```{r,echo = FALSE, fig.width=8, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$Intensity_MeanIntensity_Corr_GLUT))

ggplot(data,aes(x= Disease_state, y= Intensity_MeanIntensity_Corr_GLUT, group = Disease_state, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results/figures/data_mean-intensity_violin-plot_Disease-state.svg", width = 8, height = 6, units = "cm")
```


## Intensity cytoplasm ring 10 {.tabset} 
### each sample
```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$RadialDistribution_FracAtD_Corr_GLUT_10of10))

ggplot(data,aes(x= Metadata_SampleID, y= RadialDistribution_FracAtD_Corr_GLUT_10of10, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results/figures/data_FracAtD10_violin-plot_patients.svg", width = 20, height = 6, units = "cm")
```

### disease-state
```{r,echo = FALSE, fig.width=8, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$RadialDistribution_FracAtD_Corr_GLUT_10of10))

ggplot(data,aes(x= Disease_state, y= RadialDistribution_FracAtD_Corr_GLUT_10of10, group = Disease_state, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results/figures/data_FracAtD10_violin-plot_Disease-state.svg", width = 8, height = 6, units = "cm")
```

## Intensity cytoplasm ring 9 {.tabset} 
### each sample
```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$RadialDistribution_FracAtD_Corr_GLUT_9of10))

ggplot(data,aes(x= Metadata_SampleID, y= RadialDistribution_FracAtD_Corr_GLUT_9of10, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results/figures/data_FracAtD9_violin-plot_patients.svg", width = 20, height = 6, units = "cm")
```

### disease-state
```{r,echo = FALSE, fig.width=8, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$RadialDistribution_FracAtD_Corr_GLUT_9of10))

ggplot(data,aes(x= Disease_state, y= RadialDistribution_FracAtD_Corr_GLUT_9of10, group = Disease_state, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results/figures/data_FracAtD9_violin-plot_Disease-state.svg", width = 8, height = 6, units = "cm")
```

## Mean intensity cytoplasm ring 10 {.tabset} 
### each sample
```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$RadialDistribution_MeanFrac_Corr_GLUT_10of10))

ggplot(data,aes(x= Metadata_SampleID, y= RadialDistribution_MeanFrac_Corr_GLUT_10of10, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results/figures/data_MeanFrac10_violin-plot_patients.svg", width = 20, height = 6, units = "cm")

```

### disease-state
```{r,echo = FALSE, fig.width=8, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$RadialDistribution_MeanFrac_Corr_GLUT_10of10))

ggplot(data,aes(x= Disease_state, y= RadialDistribution_MeanFrac_Corr_GLUT_10of10, group = Disease_state, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results/figures/data_MeanFrac10_violin-plot_Disease-state.svg", width = 8, height = 6, units = "cm")

```

## Mean intensity cytoplasm ring 9 {.tabset} 
### each sample
```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$RadialDistribution_MeanFrac_Corr_GLUT_9of10))

ggplot(data,aes(x= Metadata_SampleID, y= RadialDistribution_MeanFrac_Corr_GLUT_9of10, group = Metadata_SampleID, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results/figures/data_MeanFrac9_violin-plot_patients.svg", width = 20, height = 6, units = "cm")

```

### disease-state
```{r,echo = FALSE, fig.width=8, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}
yint <- median(as.numeric(data[data$Disease_state == "Ctrl",]$RadialDistribution_MeanFrac_Corr_GLUT_9of10))

ggplot(data,aes(x= Disease_state, y= RadialDistribution_MeanFrac_Corr_GLUT_9of10, group = Disease_state, color = Disease_state, fill = Disease_state) ) +
  geom_violin(trim=FALSE) + scale_color_manual(values=c("darkblue" , "red")) + 
  scale_fill_manual(values=alpha(c("darkblue" , "red"), 0.1)) +
  geom_boxplot(width=0.1)+
  theme_classic() + geom_hline(yintercept = yint)
ggsave(filename = "results/figures/data_MeanFrac9_violin-plot_Disease-state.svg", width = 8, height = 6, units = "cm")

```


# Statistical testing using linear mixed effects models {.tabset} 

Nested approach ("GLUT intensity" ~  Disease_state + (1 | Disease_state:Metadata_SampleID)) to compensate for dependencies within the groups.

## Cytoplasmic area

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

mod0 <- lme4::lmer(AreaShape_Area ~  Disease_state + (1 | Disease_state:Metadata_SampleID), data = data, REML = FALSE); summary(mod0)
Anova(mod0)
```

## Mean intensity 

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

mod0 <- lme4::lmer(Intensity_MeanIntensity_Corr_GLUT ~  Disease_state + (1 | Disease_state:Metadata_SampleID), data = data, REML = FALSE); summary(mod0)
Anova(mod0)
```

## Intensity cytoplasm ring 10

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

mod0 <- lme4::lmer(RadialDistribution_FracAtD_Corr_GLUT_10of10 ~  Disease_state + (1 | Disease_state:Metadata_SampleID), data = data, REML = FALSE); summary(mod0)
Anova(mod0)
```

## Intensity cytoplasm ring 9

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

mod0 <- lme4::lmer(RadialDistribution_FracAtD_Corr_GLUT_9of10 ~  Disease_state + (1 | Disease_state:Metadata_SampleID), data = data, REML = FALSE); summary(mod0)
Anova(mod0)
```

## Mean intensity cytoplasm ring 10

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

mod0 <- lme4::lmer(RadialDistribution_MeanFrac_Corr_GLUT_10of10 ~  Disease_state + (1 | Disease_state:Metadata_SampleID), data = data, REML = FALSE); summary(mod0)
Anova(mod0)

```

## Mean intensity cytoplasm ring 9

```{r,echo = FALSE, fig.width=20, fig.height=6, dev = c("CairoJPEG", "cairo_pdf", "CairoSVG")}

mod0 <- lme4::lmer(RadialDistribution_MeanFrac_Corr_GLUT_9of10 ~  Disease_state + (1 | Disease_state:Metadata_SampleID), data = data, REML = FALSE); summary(mod0)
Anova(mod0)

```

