---
VignetteBuilder: knitr
Suggests: BiocStyle, knitr, rmarkdown
title: "Report non-targeted metabolomics analysis, clustering and differential abundance analysis"
author: "Sebastian Schmidt"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmarkdown::html_document:
    theme: cosmo
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: TRUE
---

```{r setup, include=FALSE, echo=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'N:/AG_Wurst/blackboard_neurodegeneration/Sebastian/Sicherung/Intenso Backup/Arbeit_Coronashutdown/non-targeted Metabolomics')
#load libraries

library(uuid)
library(extrafont)
library(readxl)
library(devEMF)
library(ggplot2)
library(lattice)
library(tidyverse)
library(plotrix)
library(fdrtool)
library(uuid)
library(devEMF)
library(EnhancedVolcano)
library(devtools)
library(ggbiplot)
library(data.table)
library(sm)
library(data.table)
library(gridExtra)
library(kableExtra)
library(DT)
library(RColorBrewer)
library(pheatmap)
```

# Loading and processing of data

***

## Count matrix

First, we load and filter the count matrix.

```{r, echo = FALSE, include = FALSE, warning = FALSE}
#############################################
## Load count matrix and tidy it

# read count matrix
sample_metabolites <- readr::read_delim("list_norm_median_cellcount_newnames.txt", delim = "\t")

# assign rownames
metabolites_filtered <- sample_metabolites %>% remove_rownames %>% column_to_rownames(var="CHEMICAL ID")
drops <- c("PATHWAY SORTORDER","SUPER PATHWAY","SUB PATHWAY", "BIOCHEMICAL", "CLIENT IDENTIFIER", "PUBCHEM", "CAS", "KEGG", "HMDB")
metabolites_filtered <- metabolites_filtered[ , !(names(metabolites_filtered) %in% drops)]

# exclude rarely detected metabolites
metabolites_filtered <- na_if(metabolites_filtered, 0)
metabolites_filtered<- metabolites_filtered[rowMeans(metabolites_filtered, na.rm = TRUE) != 0, ]
metabolites_filtered <- metabolites_filtered[rowSums(is.na(metabolites_filtered)) < 22, ] ##only metabolites with less than 30 % NA`s


```



```{r, echo = FALSE, include = TRUE, warning = FALSE}
datatable(as.data.frame(sample_metabolites), extensions = c('Buttons', 'FixedColumns'), options = list(dom = 'Bfrtip', buttons = c('csv', 'excel'), autoWidth = TRUE, scrollX = TRUE, fixedColumns = list(rightColumns = 1)))
```
## Check distribution
Log transformation increases normal distribution of sample measurements for the respective metabolites 
```{r, echo = FALSE, include = FALSE, warning = FALSE}
#############################################
## Check for normal distribution
chemical_id <- row.names(metabolites_filtered)
pdf(file = "histos_metabolites.pdf")
for (g in chemical_id){

  par(mfrow = c(1,2))
  hist(as.numeric(metabolites_filtered[c(g),]), breaks = 20, main = as.character(g))
  hist(as.numeric(log(metabolites_filtered[c(g),],2)), breaks = 20, main = "" )
}
dev.off()

statistics <- data.frame(row.names(metabolites_filtered))
names(statistics) <- "CHEMICAL ID"
statistics$shapiro <- apply(metabolites_filtered,1,function(x){shapiro.test(x[1:72])$p.value})
statistics$shapiro.log <- apply(metabolites_filtered,1,function(x){shapiro.test(log(x[1:72],2))$p.value})

## log transformation makes sense

metabolites_filtered <- log(metabolites_filtered,2)


```


```{r, echo = FALSE, include = TRUE, warning = FALSE}
datatable(as.data.frame(statistics), extensions = c('Buttons', 'FixedColumns'), options = list(dom = 'Bfrtip', buttons = c('csv', 'excel'), autoWidth = TRUE, scrollX = TRUE, fixedColumns = list(rightColumns = 1)))
```

# Dimensionality reduction methods 
***

## PCA

```{r, echo = FALSE, include = TRUE, warning = FALSE, fig.cap="PCA", figures-side, fig.show="hold", out.width="30%", dev = "CairoSVG"}
#############################################
#############################################
## Dimensionality reduction methods

#PCA


metabolites_pca <- prcomp(t(na.omit(metabolites_filtered)), center = TRUE, scale = TRUE)
samples_pca <- metabolites_pca$x


samples <- c(rep("iJ2C-R1-015",6), rep("i1JF-R1-018",6), rep("iM89-R1-005",6), rep("iC99-R1-007",6), rep("iR66-R1-007",6), rep("iG3G-R1-039",6), rep("i1E4-R1-003",6), rep("iAY6-R1-003",6), rep("iPX7-R1-001",6), rep("i88H-R1-002",6), rep("iO3H-R1-005",6), rep("i82A-R1-002",6))
condition <- c(rep("sPD",6), rep("Ctrl",6), rep("sPD",18), rep("Ctrl",12), rep("sPD",18),rep("Ctrl",12))
pca_condition <- ggbiplot(metabolites_pca, choices = c(1,2), ellipse=TRUE, groups=condition, var.axes = FALSE) +
  geom_jitter(size = 2.5, aes(color = condition)) + 
  theme_classic() +
  theme(axis.text.x = element_text(color="black"),
        axis.ticks = element_line(color="black"),
        axis.text.y = element_text(color="black"))
ggsave("PCA_condition_metabolite.emf",pca_condition, device = "emf", height = 15, width = 15, units = "cm", dpi = 400)

pca_samples <- ggbiplot(metabolites_pca, choices = c(1,2), ellipse=TRUE, groups=samples, var.axes = FALSE) +
  geom_jitter(size = 2.5, aes(color = samples)) + 
  theme_classic() +
  theme(axis.text.x = element_text(color="black"),
        axis.ticks = element_line(color="black"),
        axis.text.y = element_text(color="black"))
ggsave("PCA_samples_metabolite.emf",pca_samples, device = "emf", height = 15, width = 15, units = "cm", dpi = 400)

plot(pca_condition)
plot(pca_samples)

```


# Clustering and correlation between sample {.tabset}
***
Next, we perform clustering and correlation analysis, and plot these results in a couple of heatmaps. Most of the samples looked fine, no sample was removed as outlier.

## Distance matrix, followed by clustering

```{r, echo = FALSE, include = TRUE, warning = FALSE, fig.cap="Heatmap of distances", dev = "CairoSVG"}
#############################################
#############################################
## Correlation analyses

###################################
#with distances

#calculate distances between samples
sampleDists <- dist(t(metabolites_filtered))

#tidy matrix
sampleDistMatrix <- as.matrix(sampleDists)

#make a vector with colors and then plot heatmap
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
annotation_col = data.frame(
  Samples = samples, 
  Condition = condition
)
rownames(annotation_col) = colnames(metabolites_filtered)

dist_heatmap <- pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         col=colors,
         annotation_col = annotation_col)

emf("dist_heatmap.emf", emfPlus= FALSE, coordDPI = 400, 
    width=10, height=10, family= "Arial") 
dist_heatmap
dev.off()

dist_heatmap
```

## Correlation between samples

```{r, echo = FALSE, include = TRUE, warning = FALSE, , fig.cap="Heatmap of correlations", dev = "CairoSVG"}
###################################
#with correlations

#calculate correlations between samples
sampleDists <- cor(na.omit(metabolites_filtered))

#tidy matrix
sampleDistMatrix <- as.matrix(sampleDists)


#make a vector with colors and then plot heatmap
colors <- colorRampPalette(brewer.pal(9, "Blues"))(255)
cor_heatmap <- pheatmap(sampleDistMatrix, col=colors,  annotation_col = annotation_col)

emf("cor_heatmap.emf", emfPlus= FALSE, coordDPI = 400, 
    width=10, height=10, family= "Arial") 
cor_heatmap
dev.off()

```

## Hierarchical clustering of the samples

```{r, echo = FALSE, include = TRUE, warning = FALSE, fig.cap="Hierarchical clustering", dev = "CairoSVG"}
###################################
#simple hierarchical clustering using distances
clusters <- hclust(dist(t(metabolites_filtered)))
plot(clusters, labels = colnames(metabolites_filtered))

emf("hirachical_clustering.emf", emfPlus= FALSE, coordDPI = 400, 
    width=15, height=5, family= "Arial") 
plot(clusters, labels = colnames(metabolites_filtered))
dev.off()

```

# Differential abundance analysis
***
We performed a differential abundance analysis by parametric testing (two-sided t-test). The mean value of replicates per patient-derived cell line was used.

## Average values of replicates per patient-derived cell line

```{r, echo = FALSE, include = TRUE, warning = FALSE}
colnames <- unique(sapply(strsplit(names(metabolites_filtered)[-1], "_", fixed=TRUE), "[", 1)) 
metabolites_means <- data.frame(sapply(colnames, function(x) rowMeans(metabolites_filtered [, grep(x, names(metabolites_filtered))] ))) 
colnames(metabolites_means) <- colnames

sortList <- c("i1JF-R1-018", "iG3G-R1-039", "i1E4-R1-003", "iO3H-R1-005", "i82A-R1-002", "iJ2C-R1-015", "iM89-R1-005", "iC99-R1-007", "iR66-R1-007", "iAY6-R1-003", "iPX7-R1-001", "i88H-R1-002")
  

metabolites_means <- metabolites_means[,sortList]
metabolites_means <- metabolites_means[rowSums(is.na(metabolites_means)) < 5, ] ##only metabolites with less than 30 % NA`s

metabolites_means$"mean Ctrl" =apply(metabolites_means,1,function(x){mean(x[1:5], na.rm = TRUE)})
metabolites_means$"mean sPD" =apply(metabolites_means,1,function(x){mean(x[6:12], na.rm = TRUE)})
metabolites_means$"fc" =apply(metabolites_means,1,function(x){x[14]/x[13]})

```

```{r, echo = FALSE, include = TRUE, warning = FALSE}
datatable(as.data.frame(metabolites_means), extensions = c('Buttons', 'FixedColumns'), options = list(dom = 'Bfrtip', buttons = c('csv', 'excel'), autoWidth = TRUE, scrollX = TRUE, fixedColumns = list(rightColumns = 1)))
```

## Differential abundance analysis using two-sided t-test


```{r, echo = FALSE, include = FALSE, warning = FALSE}
metabolites_means$"p-value shapiro-test"=apply(metabolites_means,1,function(x){shapiro.test(as.numeric(x[1:12]))$p.value})
metabolites_means$"p-value wilcox-test"=apply(metabolites_means,1,function(x){wilcox.test(as.numeric(x[1:5]),as.numeric(x[6:12]))$p.value})

metabolites_means <-  metabolites_means[order(metabolites_means$`p-value wilcox-test`), ]
p_metabolites <- metabolites_means$`p-value wilcox-test`
metabolites_means$"p.adj wilcox-test" <- fdrtool(p_metabolites, statistic="pvalue",
                                           plot=TRUE, color.figure=TRUE, verbose=TRUE,
                                           cutoff.method="pct0")$qval

metabolites_means$"p-value t-test"=apply(metabolites_means,1,function(x){t.test(as.numeric(x[1:5]),as.numeric(x[6:12]))$p.value})

metabolites_means <-  metabolites_means[order(metabolites_means$`p-value t-test`), ]
p_metabolites <- metabolites_means$`p-value t-test`
metabolites_means$"p.adj t-test" <- fdrtool(p_metabolites, statistic="pvalue",
                                           plot=TRUE, color.figure=TRUE, verbose=TRUE,
                                           cutoff.method="pct0")$qval



metabolites_siglist <- row.names(metabolites_means[metabolites_means$`p.adj t-test` < 0.05,])  
  
`CHEMICAL ID` <- rownames(metabolites_means)
rownames(metabolites_means) <- NULL
metabolites_means <- cbind(`CHEMICAL ID`,metabolites_means)  


keep <- c("PATHWAY SORTORDER","SUPER PATHWAY","SUB PATHWAY", "BIOCHEMICAL", "CLIENT IDENTIFIER", "CHEMICAL ID", "HMDB", "KEGG", "HMDB", "PUBCHEM")
metabolites_ids <- sample_metabolites[ (sample_metabolites$`CHEMICAL ID` %in% metabolites_means$`CHEMICAL ID`),]
metabolites_ids <- metabolites_ids[ , (names(metabolites_ids) %in% keep)]
metabolites_means <- merge(metabolites_means, metabolites_ids, by =  "CHEMICAL ID")
metabolites_means <-  metabolites_means[order(metabolites_means$`p.adj t-test`), ]

```


And these are the results of our differential abundance analysis. Only genes with a signficant adjusted p-value are shown.


```{r, echo = FALSE, include = TRUE, warning = FALSE, message=FALSE}

metabolites_sig <- metabolites_means[metabolites_means$`p.adj t-test` < 0.05,]

datatable(metabolites_sig, extensions = c('Buttons', 'FixedColumns'), options = list(dom = 'Bfrtip', buttons = c('csv', 'excel'), autoWidth = TRUE, scrollX = TRUE, fixedColumns = list(rightColumns = 1)))

#write.table(diff_res, file = "results/tables/non-targeted-metabolomics_diff_res.csv", sep= ",", row.names = FALSE, quote = FALSE, na = "")
```


# Visualization of significantly altered metabolites {.tabset .tabset-fade .tabset-pills}

## Draw boxplots of significantly altered metabolites
```{r, echo = FALSE, include = TRUE, warning = FALSE, dev = "CairoSVG"}
metabolites_draw <- metabolites_means[,1:13]
row.names(metabolites_draw) <- metabolites_draw$`CHEMICAL ID`
metabolites_draw <- metabolites_draw[,2:13]

metabolites_draw <- t(metabolites_draw)
Condition <- c(rep("Ctrl", 5), rep("sPD", 7))
Condition_list <- data.frame(c(row.names(metabolites_draw)),Condition, row.names = TRUE )

Sample <- rownames(metabolites_draw)
metabolites_draw <- cbind(Sample, Condition_list, metabolites_draw)
rownames(metabolites_draw) <- NULL


for (m in metabolites_siglist){
  
    
  f = ggplot(metabolites_draw, aes(Condition, metabolites_draw[,m])) + 
    ggtitle(paste('chemID_',m,sep="")) + xlab("") + ylab("Raw area counts") + 
    geom_boxplot(aes(color = Condition)) +
    geom_jitter(aes(color = Condition), size = 3, width = 0.15) + 
    scale_color_manual(values=c("darkblue", "darkred")) +
    theme_classic() +
    theme(legend.position = "none") +
    theme(axis.text.x = element_text(color="black"),
          axis.ticks = element_line(color="black"),
          axis.text.y = element_text(color="black"))
  fname = paste('metabolite_chemicalID_',m,'.',sep="") 
  plot(f)
  ggsave(paste("metabolite_plots/",fname,'pdf',sep=""),f, device = "pdf", height = 7, width = 5, units = "cm", dpi = 400)
   
}
```

## Vulcano-plot
```{r, echo = FALSE, include = TRUE, warning = FALSE, dev = "CairoSVG"}

metabolites_means[,"fc"] <- log(metabolites_means[,"fc"],2)

label <- c("100001461", "1083", "330", "391", "344")



Volcano<- EnhancedVolcano(toptable = metabolites_means,
                lab = metabolites_means$`CHEMICAL ID`, x = "fc", 
                y = "p-value t-test", 
                ylim = c(0,4), xlim = c(-1,1),
                pCutoff = 0.0137,
                FCcutoff = 0.26,
                pointSize = 2.0,
                gridlines.major = FALSE,
                colConnectors = "black",
                selectLab = label,
                labCol = "black",
                drawConnectors = TRUE
)
  


Volcano

emf("metabolites_Volcano_width7.emf", emfPlus= FALSE, coordDPI = 400, 
    width=7, height=10, family= "Arial") 
Volcano
dev.off()
emf("metabolites_Volcano_width10.emf", emfPlus= FALSE, coordDPI = 400, 
    width=10, height=10, family= "Arial") 
Volcano
dev.off()
```

