---
title: "Report proteomics analysis, Pathway enrichment analysis based on all DEPs"
author: "Sebastian Schmidt"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmarkdown::html_document:
    theme: cosmo
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'F:/PD_specific_aging_signature_after_long_term_in_vitro_cultivation/Proteome')

library(knitr)
library(readxl)
library(clusterProfiler)
library(enrichplot)
library(devEMF)
library(extrafont)
library(ggplot2)
library(org.Hs.eg.db)
library(rmarkdown)
library(DT)
library(ggplot2)
library(RColorBrewer)
library(gridExtra)
library(kableExtra)
library(biomaRt)
library(ggnewscale)


```

# Summary data

Here is a summary of the DEGs 

```{r, echo = FALSE}
d <- read.delim("results/tables/proteome_diff_res_DESeq2.csv", sep = ",")
gene <- d$genes


datatable(as.data.frame(d), extensions = c('Buttons', 'FixedColumns'), options = list(dom = 'Bfrtip', buttons = c('csv', 'excel'), autoWidth = TRUE, scrollX = TRUE, fixedColumns = list(rightColumns = 1)))

```



# Enriched KEGG terms

Gene identifier changed from HGNC symbol to EntrezID
                
```{r,echo = FALSE, include=FALSE}
hs <- org.Hs.eg.db
test <- select(hs, 
       keys = gene,
       columns = c("ENTREZID"),
       keytype = "SYMBOL")
geneEntrez <- test$ENTREZID
ekegg <- enrichKEGG(gene         = geneEntrez,
                    pvalueCutoff = 0.05)



```

Here are the enriched KEGG terms

## Table
```{r,echo = FALSE}
datatable(as.data.frame(ekegg), extensions = c('Buttons', 'FixedColumns'), options = list(dom = 'Bfrtip', buttons = c('csv', 'excel'), autoWidth = TRUE, scrollX = TRUE, fixedColumns = list(rightColumns = 1)))
```

## Dotplot
```{r,echo = FALSE, fig.width=12, fig.height=15}
dotplot(ekegg, showCategory = 30)
```

## Treeplot

```{r,echo = FALSE, fig.width=12, fig.height=15}

print(try(treeplot(pairwise_termsim(ekegg), showCategory = 20, nCluster = 6)))

emf("results/figures/eKEGG2.emf",width = 10, height = 8, coordDPI = 400, family = "Arial") 
print(try(treeplot(pairwise_termsim(ekegg), showCategory = 20, nCluster = 6)))
dev.off()


```

## Emapplot

```{r,echo = FALSE, fig.width=30, fig.height=30}
emapplot(pairwise_termsim(ekegg), showCategory = 200)
```


# Enriched KEGG-Module terms

Gene identifier changed from HGNC symbol to EntrezID
                
```{r,echo = FALSE, include=FALSE}

geneEntrez <- test$ENTREZID
eMkegg <- enrichMKEGG(gene         = geneEntrez,
                    pvalueCutoff = 1,
                    qvalueCutoff = 1)

```

Here are the enriched KEGG-Module terms

## Table
```{r,echo = FALSE}
datatable(as.data.frame(eMkegg), extensions = c('Buttons', 'FixedColumns'), options = list(dom = 'Bfrtip', buttons = c('csv', 'excel'), autoWidth = TRUE, scrollX = TRUE, fixedColumns = list(rightColumns = 1)))
```

## Dotplot
```{r,echo = FALSE, fig.width=12, fig.height=15}
#dotplot(ekegg, showCategory = 30)
```

## Treeplot

```{r,echo = FALSE, fig.width=12, fig.height=15}

print(try(treeplot(pairwise_termsim(eMkegg), showCategory = 30, nCluster = 5)))

```

## Emapplot

```{r,echo = FALSE, fig.width=30, fig.height=30}
#emapplot(pairwise_termsim(ekegg), showCategory = 200)
```


# Enriched Wikipathway terms

Gene identifier changed from HGNC symbol to EntrezID

```{r,echo = FALSE, include=FALSE}
ewp <- enrichWP(geneEntrez, organism = "Homo sapiens")
```
Here are the enriched WP terms

## Table
```{r,echo = FALSE}
datatable(as.data.frame(ewp), extensions = c('Buttons', 'FixedColumns'), options = list(dom = 'Bfrtip', buttons = c('csv', 'excel'), autoWidth = TRUE, scrollX = TRUE, fixedColumns = list(rightColumns = 1)))
```

## Dotplot
```{r,echo = FALSE, fig.width=12, fig.height=5}
dotplot(ewp, showCategory = 30)
```

## Treeplot
```{r,echo = FALSE, fig.width=12, fig.height=5}
print(try(treeplot(pairwise_termsim(ewp),showCategory = 20, nCluster = 8)))
emf("results/figures/eWP.emf",width = 10, height = 8, coordDPI = 400, family = "Arial") 
print(try(treeplot(pairwise_termsim(ewp),showCategory = 20, nCluster = 8)))
dev.off()

```

## Emapplot
```{r,echo = FALSE, fig.width=30, fig.height=30}
emapplot(pairwise_termsim(ewp), showCategory = 200)
```

# Enriched Reactome terms

Gene identifier changed from HGNC symbol to EntrezID

```{r,echo = FALSE, include = FALSE, include=FALSE}
library(ReactomePA)

ereactome <- enrichPathway(gene=geneEntrez, pvalueCutoff = 0.05, readable=TRUE)
```

Here are the enriched Reactome terms

## Table
```{r,echo = FALSE}
datatable(as.data.frame(ereactome), extensions = c('Buttons', 'FixedColumns'), options = list(dom = 'Bfrtip', buttons = c('csv', 'excel'), autoWidth = TRUE, scrollX = TRUE, fixedColumns = list(rightColumns = 1)))
```

## Dotplot
```{r,echo = FALSE, fig.width=12, fig.height=10}
dotplot(ereactome, showCategory = 30)
```

## Treeplot
```{r,echo = FALSE, fig.width=15, fig.height=10}
treeplot(pairwise_termsim(ereactome), showCategory = 25, nCluster = 7)
emf("results/figures/eReactome.emf",width = 10, height = 12, coordDPI = 400, family = "Arial") 
treeplot(pairwise_termsim(ereactome), showCategory = 25, nCluster = 7)
dev.off()
```

## Emapplot
```{r,echo = FALSE, fig.width=30, fig.height=30}
emapplot(pairwise_termsim(ereactome), showCategory = 200)
```

